\documentclass{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{geometry}

\geometry{margin=1in}

\title{Lecture Summary: Determinants (Part 1)}
\author{}
\date{}

\begin{document}

\maketitle

\section*{Source: Lec 19.pdf}

\section*{Key Points}

\begin{itemize}
  \item \textbf{Definition of Determinants:}
    \begin{itemize}
      \item The determinant is a scalar value associated with a square matrix.
      \item Notation: $\det(A)$ or $|A|$.
      \item Applications:
        \begin{itemize}
          \item Solving systems of linear equations.
          \item Finding the inverse of a matrix.
          \item Applications in calculus and other advanced topics.
        \end{itemize}
    \end{itemize}

  \item \textbf{Determinants of Small Matrices:}
    \begin{itemize}
      \item \textbf{1x1 Matrix:} For $A = [a]$, $\det(A) = a$.
      \item \textbf{2x2 Matrix:}
        \[
          A =
          \begin{bmatrix}
            a & b \\
            c & d
          \end{bmatrix},
          \quad
          \det(A) = ad - bc.
        \]
      \item \textbf{3x3 Matrix:}
        \[
          A =
          \begin{bmatrix}
            a_{11} & a_{12} & a_{13} \\
            a_{21} & a_{22} & a_{23} \\
            a_{31} & a_{32} & a_{33}
          \end{bmatrix},
        \]
        \[
          \det(A) = a_{11}
          \begin{vmatrix}
            a_{22} & a_{23} \\
            a_{32} & a_{33}
          \end{vmatrix}
          - a_{12}
          \begin{vmatrix}
            a_{21} & a_{23} \\
            a_{31} & a_{33}
          \end{vmatrix}
          + a_{13}
          \begin{vmatrix}
            a_{21} & a_{22} \\
            a_{31} & a_{32}
          \end{vmatrix}.
        \]
    \end{itemize}

  \item \textbf{Properties of Determinants:}
    \begin{itemize}
      \item \textbf{Multiplication:} $\det(AB) = \det(A) \cdot \det(B)$.
      \item \textbf{Inverse:} $\det(A^{-1}) = \frac{1}{\det(A)}$.
      \item \textbf{Row/Column Operations:}
        \begin{itemize}
          \item Swapping two rows or columns changes the determinant's sign.
          \item Adding a multiple of one row (or column) to another does not change the determinant.
          \item Multiplying all entries of a row or column by a constant $t$ scales the determinant by $t$.
        \end{itemize}
    \end{itemize}
\end{itemize}

\section*{Simplified Explanation}

\textbf{Example: Determinant of a 2x2 Matrix}
For $A =
\begin{bmatrix}
  2 & 3 \\
  6 & 10
\end{bmatrix}$,
\[
  \det(A) = (2 \cdot 10) - (3 \cdot 6) = 20 - 18 = 2.
\]

\textbf{Example: Determinant of a 3x3 Matrix}
For $A =
\begin{bmatrix}
  2 & 4 & 1 \\
  3 & 8 & 7 \\
  5 & 6 & 9
\end{bmatrix}$:
\[
  \det(A) = 2 \cdot
  \begin{vmatrix}
    8 & 7 \\
    6 & 9
  \end{vmatrix}
  - 4 \cdot
  \begin{vmatrix}
    3 & 7 \\
    5 & 9
  \end{vmatrix}
  + 1 \cdot
  \begin{vmatrix}
    3 & 8 \\
    5 & 6
  \end{vmatrix}.
\]
\[
  \det(A) = 2 \cdot (72 - 42) - 4 \cdot (27 - 35) + 1 \cdot (18 - 40).
\]
\[
  \det(A) = 2 \cdot 30 - 4 \cdot (-8) + 1 \cdot (-22) = 70.
\]

\section*{Special Matrices and Their Determinants}

\begin{itemize}
  \item \textbf{Identity Matrix:}
    For $I_2 =
    \begin{bmatrix}
      1 & 0 \\
      0 & 1
    \end{bmatrix}$, $\det(I_2) = 1$.
    Similarly, for $I_3 =
    \begin{bmatrix}
      1 & 0 & 0 \\
      0 & 1 & 0 \\
      0 & 0 & 1
    \end{bmatrix}$, $\det(I_3) = 1$.
\end{itemize}

\section*{Conclusion}

In this lecture, we explored:
\begin{itemize}
  \item The concept of determinants for 1x1, 2x2, and 3x3 matrices.
  \item Properties of determinants, including their behavior under multiplication, inversion, and row/column operations.
  \item Determinants are computed inductively, with larger matrices reduced to smaller ones.
\end{itemize}

These foundational properties of determinants will be critical in solving linear equations, understanding matrix behavior, and various applications in higher mathematics.

\end{document}
