\documentclass{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{geometry}

\geometry{margin=1in}

\title{Lecture Summary: Solution to a System of Linear Equations with an Invertible Coefficient Matrix}
\author{}
\date{}

\begin{document}

\maketitle

\section*{Source: Lec23.pdf}

\section*{Key Points}

\begin{itemize}
  \item \textbf{Invertibility of Square Matrices:}
    \begin{itemize}
      \item A square matrix $A$ is invertible if there exists a matrix $A^{-1}$ such that:
        \[
          A \cdot A^{-1} = A^{-1} \cdot A = I,
        \]
        where $I$ is the identity matrix.
      \item Invertibility is equivalent to $\det(A) \neq 0$. If $\det(A) = 0$, $A$ is not invertible.
    \end{itemize}

  \item \textbf{Inverse of a $2 \times 2$ Matrix:}
    \begin{itemize}
      \item For $A =
        \begin{bmatrix}
          a & b \\
          c & d
        \end{bmatrix}$:
        \[
          A^{-1} = \frac{1}{\det(A)}
          \begin{bmatrix}
            d & -b \\
            -c & a
          \end{bmatrix}, \quad \text{where } \det(A) = ad - bc.
        \]
    \end{itemize}

  \item \textbf{Adjugate Matrix:}
    \begin{itemize}
      \item The adjugate (or adjoint) matrix, $\text{adj}(A)$, is the transpose of the cofactor matrix of $A$.
      \item If $A$ is invertible:
        \[
          A^{-1} = \frac{1}{\det(A)} \cdot \text{adj}(A).
        \]
    \end{itemize}

  \item \textbf{Solution of a Linear System ($Ax = b$):}
    \begin{itemize}
      \item If $A$ is invertible, the solution is:
        \[
          x = A^{-1}b.
        \]
    \end{itemize}

  \item \textbf{Homogeneous Systems:}
    \begin{itemize}
      \item A system is homogeneous if $Ax = 0$.
      \item For a square matrix $A$:
        \begin{enumerate}
          \item If $\det(A) \neq 0$, the only solution is $x = 0$.
          \item If $\det(A) = 0$, the system has infinitely many solutions.
        \end{enumerate}
    \end{itemize}
\end{itemize}

\section*{Simplified Explanation}

\textbf{Example 1: Solving a System Using Inverses}
Given the system:
\[
  8x_1 + 8x_2 + 4x_3 = 1960, \quad 12x_1 + 5x_2 + 7x_3 = 2215, \quad 3x_1 + 2x_2 + 5x_3 = 1135,
\]
matrix representation:
\[
  A =
  \begin{bmatrix}
    8 & 8 & 4 \\
    12 & 5 & 7 \\
    3 & 2 & 5
  \end{bmatrix},
  \quad
  b =
  \begin{bmatrix}
    1960 \\
    2215 \\
    1135
  \end{bmatrix}.
\]
Compute $\det(A)$:
\[
  \det(A) = -188 \neq 0 \quad \Rightarrow A \text{ is invertible}.
\]
Using:
\[
  A^{-1} = \frac{1}{\det(A)} \cdot \text{adj}(A),
\]
and multiplying $A^{-1}b$, the solution is:
\[
  x_1 = 45, \quad x_2 = 125, \quad x_3 = 150.
\]

\textbf{Example 2: Homogeneous System}
If $Ax = 0$:
\begin{itemize}
  \item If $\det(A) \neq 0$, $x = 0$ is the only solution.
  \item If $\det(A) = 0$, there are infinitely many solutions.
\end{itemize}

\section*{Conclusion}

In this lecture, we:
\begin{itemize}
  \item Explored the relationship between the determinant and invertibility of square matrices.
  \item Derived the formula for $A^{-1}$ using the adjugate matrix and applied it to solve linear systems.
  \item Introduced homogeneous systems and their solution behavior based on $\det(A)$.
\end{itemize}

These concepts are foundational for solving linear equations and understanding matrix properties in linear algebra.

\end{document}
